|

### 2021-08-14 appendix/appendix.Rmd
appendix fix model primitives

# A Model of Legislature Size and Service Provision

## Full model

Consider an strategic interaction comprised of a mayor ($M$), and $N \geq 3$ of city councilors, indexed in $i \in \{1, 2, \cdots, N \}$. The mayor is head of the local Executive, and has the prerogative of proposing a vector of policies, that is voted by the city councilors. The policy proposal, if accepted, is implemented.

Every city councilor has a party affiliation, and the party label may or may not match the mayor's affiliation. If the politician shares the same party as the mayor, we say that it belongs to the government coalition ($G \subset \{1, 2, \cdots, N \}$). Otherwise, we assume that the politician belongs to the opposition $O \not\subset G$. The party affiliation is important because it provides a political benefit or cost for the councilor. If the mayor implements the policy, then a councilor aligned with the mayor receives a policy benefit $p > 0$. If the politician belongs to the opposition, then the mayoral policies impose her a cost of $-p$. For simplicity, we assume that these costs are symmetric.

However, if the mayor's policy proposal lose majoritarian support from the council, then we assume that a reversal policy takes place, where the resources diminish by a factor of $\delta \in (0,1)$, and a random councilor is recognized to make a proposal.\footnote{In a model similar to [Baron Ferejohn / Cliff XXXX].} The councilor proposes the new policy vector, and the council votes. If the policy is rejected, then another councilor is recognized to make the offer, and this goes on until one offer is finally accepted by the majority of the legislators. For simplicity, we assume that the councilor's offers generate no political benefits or costs, so that if a councilor proposes the policy vector, then $p=0$. This may sound restrictive, but we will show that councilors propose zero public goods in equilibrium, as they have only to please their particularistic constituencies to get reelected.

To convince the councilors, the politician making the proposal has three vectors of policies at her disposal. First, the proposer may give targeted goods for the councilor $i$ that she is trying to earn support, $t_i \geq 0$. These targeted goods improve the reelection rates of the councilor. Second, the proposer may also give private goods $x_i \geq 0$ to a councilor $i$. Private goods do not affect the reelection rates of the politicians, but provide instant benefits for the receiver. Finally, the proposer can provide public goods $g \geq 0$. Public goods help the mayor to get reelected, and also affect the policy benefits received by the politicians, depending on the party affiliation.

For the mayor, the expected utility function is a sum of the utility from rents and benefits from reelection. Both the utilities from rents ($u(r)$) and the probability of getting reelected $\pi(g)$ are concave functions, meaning that more rents or public goods increase the utility at a decreasing rate. The probability of getting reelected is also multiplied by the benefit from holding office $B_M > 0$. This term captures the tangible and intangible gains that the mayor perceives from being reelected into the position. The utility is multiplied by an indicator $\mathbb{I}$ that captures whether the mayor's proposal was accepted by the city council.\footnote{The indicator function is defined as
\[
I(condition) \ = \ \twopdo{1}{\text{Condition is satisfied.}}{0}.
\]
} The expected utility for the mayor is:

\[
\mathbb{E}U(r, g, \bf{x}, \bf{t}) \ = \ \mathbb{I}(\text{Approval})(u(r) + B_M\pi(g))
\]

And the mayor chooses the optimal policy subject to a budget constraint. The municipality has $R$ resources, and we assume that the municipality cannot run debts. This budget balance constraint requires the policy vector $(r, g, \bf{x}, \bf{t})$ to satisfy the following rule:

\[
r + g + \sum_{i=1}^{N}t_i + \sum_{i=1}^{N} x_i \ \leq \ \delta^kR
\]

And the mayor's proposal happens when $k=0$. This makes mayoral budget inequality to satisfy: $r + g + \sum_{i=1}^{N}t_i + \sum_{i=1}^{N} x_i \ \leq \ R$. Otherwise, in the reversal proposals, the resources diminish by a $\delta^k$ rate, with $k$ being the number of unsuccessful offers by the councilors.

We can decompose the councilor's utility function into three major components. First, the political gains or losses (depending on the councilor's party affiliation); second, the rents $x_i$ offered by the policy proposer to the councilor; and finally, the transfers for the constituents $t_i$, offered by the policy proposer to the councilor. As the constituent transfer $t_i$ affects chance of winning the election $\pi_C(t_i)$, we multiply this chance by the benefits accrued from remaining in office, $B > 0$. Again, as in the mayor's case, the utility is only realized when the city council accepts the policy proposed, so that we multiply the approval indicator $\mathbb{I}(.)$. The councilor's expected utility is equal to:

\[
\mathbb{E}U(r, g, \bf{x}, \bf{t}) \ = \  \mathbb{I}(\text{Approval})((-1)^{\mathbb{I}(i \in O)}p + x_i + B\pi_C(t_i))
\]

For simplicity, we assume that the councilor's probability of reelection satisfy a simple cutoff rule:

\[
\pi_C(t_i) \ = \ \twopdo{1}{t_i \geq \underline{t}}{0}
\]

This assumption simplifies the calculations, and it is reasonable as many models have the voter's decision as a cutoff strategy [@XXXX]. We also assume, for convenience, that $B > \underline{t}$. Otherwise, the councilor would prefer to pocket the targeted goods, and not being reelected.

@@

\begin{enumerate}
  \item The mayor learns how many government $|G|$ and opposition $|O|$ legislators were elected.
  \item The mayor proposes a policy vector $(r, g, \bf{x}, \bf{t})_M$.
  \item The city council votes the proposal.
  \begin{itemize}
    \item If the council accepts, then the policy is implemented, and the game ends. 
    \item Otherwise, the reversal policy is implemented.
  \end{itemize}
  \item (Reversal Policy) At the beginning of each iteration of the reversal policy, one councilor is randomly recognized to be the proposer. She proposes a vector $(\bf{x}, \bf{t})$, with $r =  g = 0$. If the proposal is accepted, it is implemented. If the proposal is rejected, a new round of reversal start, but with the municipal resources reduced by a factor $\delta$.
\end{enumerate}

This is a bargaining game, along the lines of @rubinsteinXXXX. For political science models similar to this model, see @baronfereXXXX and @cliffsXXXX.

The solution for this game consists in finding a Stationary Sub-game Perfect Nash Equilibrium using backward induction. Although the game tree may extend infinitely, a Stationary Sub-game Perfect Equilibrium assumes that if a politician accepts an offer at time $t+1$, she should accept the offer at time $t$. This gives us an objective of finding the offer that will be accepted at any point in the game. Moreover, applying this reasoning we can find the optimal offer at $t=0$, which represents no delay in policy implementation. The mayor calibrates the offer for approval offering the benefits that councilors would receive, should they reject the offer and enter the reversal stage.

\begin{definition}
The solution for the game consists in a pair of strategic policy vectors for the mayor, and the city councilors. The mayor's policy vector is comprised of a public goods provision $g^*$ and XXXX
\end{definition}

Using backward induction, suppose that there were $k$ prior rejections, and the municipal resources shrank by $\delta^{k}R$. At this stage, the politicians are similar with regard to political affiliation, so that we drop the party label. A given councilor is recognized as the proposer with $1/N$ probability. The recognized councilor has to convince at least one-half of the council members to agree with her policy proposal. Therefore, a legislator $i$, that receives an offer from a proposer ($x_i, t_i$), at round $k$, will accept the offer if it is at least as good as waiting for the next round.

\[
x_i + B\pi_C(t_i) \ \geq \ \dfrac{1}{N}\left[\delta^{k+1}R - \dfrac{N}{2}(x_{-i} + t_{-i}) + B\pi_C(t_i)\right] + \dfrac{1}{2}(x_i + B\pi_C(t_i))
\]

At the $k$-th stage, the players are equivalent. Therefore, the offer $x_i, t_i$, and the offer for the other players, ($x_{-i} + t_{-i}$), are the same. Moreover, as $B > \underline{t}$, $t^* = t_i = t_{-i} = \underline{t}$, and the offer that would make legislator $i$ accept without further delay is:

\[
x + B \ \geq \ \dfrac{1}{N}\left(\delta^{k+1}R - \dfrac{N}{2}(x + \underline{t}) + B\right) + \dfrac{1}{2}(x + B)
\]

And after some algebra, the optimal private goods offer has to satisfy:

\[
x \geq \dfrac{1}{N}\left( \delta^{k+1}R + B \right) - \dfrac{1}{2}(B + \underline{t}) \equiv \underline{x}(k)
\]

For all $k \in \{1, 2, \cdots, \infty\}$.

\begin{lemma}
Suppose that the council rejected the mayor's offer. Then, at the $k$-th stage of the game, the proposer will offer $t^* = \underline{t}$ and $x^* = \underline{x}(k)$ for half of the members, keeping for her $t_i = \max\{0, \underline{t}\}$, and

\[
x_i \ = \ R\delta^k\left( 1 - \dfrac{\delta}{2} \right) + \dfrac{N}{4}(B-\underline{t}) + \dfrac{B}{2}
\]

Subject to the budget constraint:

\[
\dfrac{N}{2}\left(\underline{t}+\underline{x}(k)\right) + x_i  \ \leq \ \delta^kR
\]

The sequence of strategies derived here are an equilibrium for all $k \in \{1, 2, \cdots, \infty\}$.

\end{lemma}

\begin{proof}
The proof is direct, after collecting the terms from the previous discussion.

Note that after some iterations, the budget may shrink enough so that $t^* = \underline{t}$ is unfeasible. This means that there exists a $k^*$ such that, when $k > k^*$, $t^* = 0$, and the councilors divide the remaining budget as rents. This case is trivial to solve, making $\underline{x}(k) = \dfrac{1}{N}\left(\delta^{k+1}R\right)$ and $x_i = \delta^kR\left(1 - \dfrac{\delta}{2}\right)$.
\end{proof}

We now have the entire strategy profile for the city councilors, provided that they reject the initial mayor's proposal. Proceeding by backward induction, we solve for the mayor's proposal that wins the support of the majority of the council. In the mayor's proposal policy concerns come into play and the offer for a government-aligned councilor is different than the offer for an opposition councilor. For a government aligned councilor, assuming the councilor is the pivotal for the implementation, the accepted offer $(x_m^g, t_m^g)$ has to be better than rejecting the proposal, and waiting for the next round:

\[
p + x_m^g + B\pi_C(t_m^g) \ \geq \ \dfrac{1}{N}\left( \delta R -\dfrac{N}{2}(\underline{x}(1) + \underline{t}) + B\pi_C(t_i))  \right) + \left(1 - \dfrac{1}{N}\right)\left(\dfrac{1}{2}\right)(\underline{x}(1) + B\pi_C(t_i))
\]

Note that $\underline{x}(1)$ is the offer of private goods if the councilor rejects the mayors offer, so that $k = 1$. As the benefit of retention $B$ is higher than the cost of providing the targeted goods $\underline{t}$, the optimal provision of targeted goods satisfy the threshold for reelection: $t^* = t_m^g = t_i = \underline{t}$. Rewriting the equations leads us to:

\[
p + x_m^g + B \ \geq \ \dfrac{1}{N}\left( \delta R -\dfrac{N}{2}(\underline{x}(1) + \underline{t}) + B)  \right) + \left(1 - \dfrac{1}{N}\right)\left(\dfrac{1}{2}\right)(\underline{x}(1) + B)
\]

And after some algebra:

\[
x_m^g \ \geq \ -p +\dfrac{1}{2N+1}\left[2\delta R \left(1 + \dfrac{1-\delta}{2N}\right)-N(\underline{t} + B) + B\right] \ \equiv \ \underline{x}_m^g
\]

For the opposition councilor, the offer has to compensate the political loss of having a mayor that is not aligned with the councilor's ideology. The optimal offer satisfy:

\[
-p + x_o^g + B \ \geq \ \dfrac{1}{N}\left( \delta R -\dfrac{N}{2}(\underline{x}(1) + \underline{t}) + B)  \right) + \left(1 - \dfrac{1}{N}\right)\left(\dfrac{1}{2}\right)(\underline{x}(1) + B)
\]

And after some algebra:

\[
x_o^g \ \geq \ p + \dfrac{1}{2N+1}\left[2\delta R \left(1 + \dfrac{1-\delta}{2N}\right)-N(\underline{t} + B) + B\right] \ \equiv \ \underline{x}_o^g
\]

\begin{lemma}
The offer for a government-aligned councilor is always lower than the offer for an opposition councilor.
\end{lemma}

\begin{proof}
Note that the $\underline{x}_m^g = 2p + \underline{x}_o^g$. As the political disagreement is higher than zero, $\underline{x}_m^g = - 2p + \underline{x}_m^o < \underline{x}_m^g$. Moreover, note that $\underline{x}_m^g - \underline{x}_m^o = -2p$.
\end{proof}

This characterizes the offer that the mayor will make to win the support of the council. The next Definition helps to decompose the politician's offer into two components: rents and policy compensation.

\begin{definition}
Define the rents offered from the bargaining process as

\[
b(N) \ = \ \dfrac{1}{2N+1}\left[2\delta R \left(1 + \dfrac{1-\delta}{2N}\right)-N(\underline{t} + B) + B\right]
\]

When the legislature size is $N$, the optimal offers for an opposition member is:

\[
\underline{x}_m^o(N) = b(N) + p
\]

And the optimal offer for a government member is:

\[
\underline{x}_m^g(N) = b(N) - p
\]

\end{definition}

Now, let the number of government politicians be equal to $\eta_g$, and the ex-ante probability of electing a councilor aligned with the government equal to $\gamma \in (0,1)$. We define the cost to govern $C_G(N)$ as three parts equation: first, the costs of making the electoral transfers for at least half of the councilors; second, the rents paid in bargaining; and last, the political costs, that depend on the amount of politicians that are affiliated with the mayor's party. After simplifications, the equation describing the costs of govern is as follows:

\[
C_G(N) = \dfrac{N}{2}(\underline{t} + b(N) + p) - 2p\sum_{\eta_g = 0}^{N/2}\mathbb{P}(\eta_g)\eta_g
\]

And the probability that $\eta_g$ councilors belong to the mayor's coalition is equal to $\mathbb{P}(\eta_g) = \binom{N}{\eta_g}\gamma^{\eta_g}(1-\gamma)^{N-\eta_g}$, which is the binomial probability given that the ex-ante chance to elect an aligned councilor is equal to $\gamma$.

We need now to derive the optimal rents ($r$) and public goods provision ($g$) proposed by the mayor. The mayor benefits from public goods provision, as it increases the chances of her reelection. However, she prefers to contribute as minimum as possible for the public goods, and extract the remaining resources as political rents. Therefore, her objective is to maximize the expected utility subject to the budget constraint.

@@

\begin{proof}
In order to prove this result, we need to show that the mayor's expected utility satisfy the increasing differences in $g$ and $N$. This would mean that when increasing the size of the council, the optimal solution $g^*$ would also increase. However, as $N$ increases discretely, we cannot take a derivative on $N$ and use the implicit function theorem. We use Monotone Comparative Statics [@milgromshannonXXXX] to derive these results.

@@

Therefore, to satisfy increasing differences in XXXX,

@@

This gives us our first empirical hypothesis to test: if the cost of govern decreases when the legislature increases, then the public goods provision increase when the legislature size increase. But we still need to investigate further how the costs of govern would decrease with legislature size. On the one hand, it seems evident that with the additional legislator, comes also an additional rent and a possible veto player. In the next Lemma we show that the rents always decrease when we increase the number of legislators.

\begin{lemma}\label{lemmabarg}
The amount of rents required to win support from a given legislator decreases when the number of legislators increase.
\end{lemma}

@@

The proof consists in showing that when we increase the size of the legislature from $N$ to $N+1$, the amount of rents decrease.

\[
\begin{aligned}
b(N+1) - b(N) \ = \ & \dfrac{1}{2(N+1)+1}\left[2\delta R \left(1 + \dfrac{1-\delta}{2(N+1)}\right)-(N+1)(\underline{t} + B) + B\right] \\
& - \dfrac{1}{2N+1}\left[2\delta R \left(1 + \dfrac{1-\delta}{2N}\right)-N(\underline{t} + B) + B\right] \\
= \ & \dfrac{1}{2N+1}\left[\left(1-\dfrac{2}{2N+3}\right)\left(2\delta R \left(1 + \dfrac{1-\delta}{2(N+1)}\right)-N(\underline{t} + B) + \underline{t}\right) \right. \\
& - \left. 2\delta R \left(1 + \dfrac{1-\delta}{2N}\right)-N(\underline{t} + B) + B\right] \\
= \ & -\dfrac{1}{2N+1}\left[\dfrac{2}{2N+3}\left(2\delta R \left(1 + \dfrac{1-\delta}{2(N+1)}\right)-N(\underline{t} + B) + \underline{t}\right) \right. \\
& + \left. \dfrac{2 R \delta (1-\delta)}{2N} + \underline{t} + B\right] \equiv \Delta b(N) \\
< \ &0
\end{aligned}
\]

Therefore, larger legislature sizes decrease the costs of bargaining. This because of two effects: first, the chance of being the proposer decreases; second, the chance of being selected to receive an offer also decreases.

@@

However, the decrease in rents is not sufficient to have lower costs to govern when we increase the number of legislators. The political costs, if high enough, can offset the gains with lower rents. Higher political costs result from having more members of the opposition in the legislature. The next proposition shows the sufficient condition for having a lower cost of govern when we increase the number of legislators.

\begin{proposition}\label{mainprop}

There exists an ex-ante probability of electing a councilor aligned with the government $\overline{\gamma}$ such that, for all $\gamma > \overline{\gamma}$, the cost of govern decreases as $N$ increases, i.e., $C_G(N+1) \leq C_G(N)$.

\end{proposition}

@@

Let us rewrite the results as:

\[
XXXX \geq 0
\]

Rearranging the terms:

\[
- \underline{t} - \underline{x}_o^g(N+1) - \dfrac{N}{2}(\underline{x}_o^g(N+1) - \underline{x}_o^g(N-1)) -2p \left[\sum_{\eta_g = 0}^{N/2+1}\mathbb{P}(\eta_g)\eta_g - \sum_{\eta_g = 0}^{N/2}\mathbb{P}(\eta_g)\eta_g\right]
\]

Let $\Delta\eta_n(\gamma) \equiv \left[\sum_{\eta_g = 0}^{N/2+1}\mathbb{P}(\eta_g)\eta_g - \sum_{\eta_g = 0}^{N/2}\mathbb{P}(\eta_g)\eta_g\right]$. There is no closed form solution for this equation. We can approximate the boundaries using @wXXXX but graphically, we can have a good sense of the XXXX:

XXXX GRAPH DELTA GAMMA HERE

And solving for $\Delta\eta_n(\gamma)$ we have that:

\[
XXXX
\]

The natural boundaries for $\Delta\eta_n(\gamma)$ are minus-one and one. Minus-one means that the additional legislature seat went to the opposition while one means that the additional legislature seat went to the government. XXXX

@@

Taking stocks, the model provides two testable hypothesis. Let a legislature that previously had $N$ legislators, and now increased the legislature by $N+1$. Then:

@@

## Paper Version of the Model

In the main paper, we simplify the analysis with two assumptions regarding the costs of govern. First, we assume a simpler reversal where there is only one reversal stage and after that a reversal policy of zero in all policy components take place. Second, we assume that the cost for govern will depend on the ex-ante expected number of legislator belonging to the government and the opposition. In this case, the rents $b(N)$ become:

\[
XXXX
\]

And Proposition \ref{mainprop} becomes:

\begin{proposition}

XXXX

\[
XXXX
\]

\end{proposition}

\begin{proof}
XXXX
\end{proof}

All the dynamics remain the same but the second hypothesis becomes clearer, with a simplified quantity for the probability of electing a government-aligned politician, $\overline{\gamma}$. The simplified $\overline{\gamma}$ is helpful because it shows that the political concerns require a probability larger than one-half, with a precise amount depending on the XXXX.

>>>>

## A Model of Legislature Size and Service Provision

### Primitives of the Model

Consider an strategic interaction between a mayor ($M$) and $N \geq 3$ city councilors, indexed in $i \in \{1, 2, \cdots, N \}$. The mayor is the head of the local executive branch, and has the prerogative of proposing a vector of policies, that is voted by the city councilors. Policy proposals are a combination of public goods provision $g$, rents for the mayor $r$, and a vector of transfer for city councilors $\mathbb{x}$.

The city council votes the mayor's proposal and if they accept it, it is implemented. Otherwise, a reversal policy takes place, where the decision leaves the mayor's hands and either immediately stops or transfers to the council. The types of reversal policies are crucial for our argument, as it changes the relative strength of the councilors in their bargaining with the mayor. We investigate three types of reversal policies: a \textit{non-partisan} reversal mechanism, where city councilors have no partisan concerns; a \textit{partisan} reversal mechanism, where the council is divided in support and opposition for the mayor; and finally a \textit{hybrid} reversal mechanism, that combines partisan concerns with non-partisan rent extraction. In the main paper, we report only the hybrid model, as it is the more realistic scenario. Despite that, comparing the partisan and non-partisan reversal mechanisms illustrate clearly our argument that partisan a mechanism with only politics (city councilors have parties, that generates political benefits and costs), and an intermediate mechanism (a more realist mechanism, including both extremes).

To govern, the mayor has to convince at least half of the city councilors to support her policy proposals, and she has to use the policy instruments at her disposal. These instruments will vary depending on the mechanism that we will investigate, but they are two in nature: targeted goods for the city councilor's constituency and private goods for the city councilor. The targeted goods improve the councilor's reelection chances while the private goods can be understood as corruption, aimed at ``buying off'' the support of a recipient councilor. The use of the policy instruments, will generate a cost of govern, $C_G$, that will vary with the policy choices and the legislature size.

Finally, the mayor has to propose a level of public goods provision ($g$) and rents ($r$). Public goods provision help the mayor to get reelected. Rents are for direct mayor's consumption, and does not contribute to the mayor's electoral success. This makes the mayor's expected utility a sum of the utility from rents and benefits from reelection. Both the utilities from rents ($u(r)$) and the probability of reelection $\pi(g)$ are concave functions, meaning that more rents or public goods increase the utility at a decreasing rate. The probability of reelection is multiplied by the benefit from holding office $B_M > 0$. This benefit captures the tangible and intangible gains that the mayor perceives from holding the public office. The utility is multiplied by an indicator $\mathbb{I}$ that captures whether the mayor's proposal was accepted by the city council.\footnote{The indicator function is defined as
\[
I(condition) \ = \ \twopdo{1}{\text{Condition is satisfied.}}{0}.
\]
} The expected utility for the mayor is:

\[
\mathbb{E}U(r, g) \ = \ \mathbb{I}(\text{Approval})(u(r) + B_M\pi(g))
\]

The policy choices of the mayor are subject to the municipal budget constraint. The municipality has $R$ resources, and we assume the municipality cannot run debts. Let $C_G$ the costs of govern. The budget balance constraint requires the offers that the mayor makes to the city councilors to satisfy the following rule:

\[
r + g + C_G \ \leq \ R
\]

The expected utility for the city councilors depend on the mechanism that will determine the reversal policy. We explore three different mechanisms. First, in the *non-partisan legislature* mechanism, we assume that when the council rejects the mayor's office, a reversal stage starts with the selection of one councilor randomly. Resources diminish by a factor of $\delta \in (0,1)$, and the selected councilor will have to make a proposal. If her proposal is accepted, it is implemented. If it is rejected, the budget shrinks again by a factor $\delta$, and another city councilor is recognized to make the proposal. The process than repeats, until one proposal is finally accepted.

In the second reversal mode, the *partisan legislature* mechanism, we assume that each councilor has a party affiliation. Party affiliations are mutually exclusive (a councilor cannot belong to two parties at once). If a city councilor belongs to a party aligned with the mayor, we say that she belongs to the government coalition ($G \subset \{1, 2, \cdots, N \}$). Otherwise, the city councilor belongs to the opposition $O \not\subset G$. The party affiliation play an essential role in the *partisan legislature* mechanism: if a city councilor that belongs to the mayor's coalition rejects the mayor's proposal, she loses the election. Whereas, if the city councilor belongs to the opposition, she loses the election by supporting mayor's proposal. Although these assumptions are strong, they capture the main underlying political dynamics. Partisanship means in most political interactions that if a city councilor aligns with the mayor, her electorate has preferences toward policies that the mayor would implement. Voting against the mayor's proposals would represent to defy the partisan brand.

In the last reversal mode, which is the reversal mechanism we use in the paper, we combine both partisanship with legislature bargaining rents. Councilors in this mechanism receive a partisan utility from the mayor's policy, but if they reject the policy, a reversal policy similar to the *non-partisan legislature* mechanism takes place. In the mayor's proposal stage, city councilors that belong to the mayor's coalition receive a gain of $p > 0$ when the mayor's policy is implemented. Opposition members, on the other hand, receive $-p$ when the mayor's proposal is implemented. If the mayor's proposal is rejected, as city councilors do not control the Executive, no public goods are provided ($p=0$). Moreover, in equilibrium, we assume that city councilors would prefer to receive rents instead of providing public goods.

@@

\begin{enumerate}
  \item The mayor learns how many government $|G|$ and opposition $|O|$ legislators were elected.
  \item The mayor proposes a policy vector $(r, g, \bf{x})_M$.
  \item The city council votes the proposal.
  \begin{itemize}
    \item If the council accepts, the policy is implemented and the game ends. 
    \item Otherwise, the reversal policy is implemented.
  \end{itemize}
  \item (Reversal Policy) Depends on which mechanism, we will have three different reversal policies. We will detail them below.
\end{enumerate}

This is a bargaining game, along the lines of @rubinstein1982perfect. For a political science model similar to this model, see @baron1989bargaining.

The solution for this game consists in finding a Sub-game Perfect Nash Equilibrium using backward induction. In the model with infinitely repeated proposals, we extend the equilibrium concept to require that the equilibrium is Stationary. A Stationary Sub-game Perfect Equilibrium is based on the assumption that if a politician accepts an offer at time $t+1$, she should accept the same offer at time $t$. This gives us an objective of finding the offer that will be accepted at any point in the game. Applying this reasoning, we can find the optimal offer at $t=0$, which represents no delay in policy implementation.

\begin{definition}
The solution for the game consists in a pair of strategic policy vectors that for the mayor, and the city councilors, indexed by the round of the implementation. The mayor's policy vector is comprised of a public goods provision $g^*$, rents $r^*$, $\mathbb{t}$ XXXX
\end{definition}

### The Mayor's Stage

Assume that we have worked out the solution using backward induction, and when there are $N$ councilors, we find that the cost of govern for the mayor is $C_G(N)$. This assumption will be dropped when we work each of the cases below.

We now derive the optimal rents ($r$) and public goods provision ($g$) proposed by the mayor. The mayor benefits from public goods provision, as it increases the chances of her reelection. However, she prefers to contribute as minimum as possible for the public goods, and extract the remaining resources as political rents. Her objective is to maximize the following expected utility, subject to the municipal budget constraint.

@@

\begin{proof}
In order to prove this result, we need to show that the mayor's expected utility satisfy the increasing differences in $g$ and $N$. This would mean that when increasing the size of the council, the optimal solution $g^*$ would also increase. However, as $N$ increases discretely, we cannot take a derivative on $N$ and use the implicit function theorem. Instead, we use monotone comparative statics [@milgrom1994monotone] to derive these results.

@@

Therefore, to satisfy increasing differences in the public goods provision and in the legislature size, the costs of govern has to decrease with legislature size.
\end{proof}

This provides us our first empirical hypothesis to test: *if the cost of govern decreases when the legislature increases, then the public goods provision increase when the legislature size increase*.

At this stage, we need to solve for the costs of govern to determine whether this hypothesis is true or not. Below we show that it is true for the *partisan* and *hybrid* mechanisms, but it is not true for the *non-partisan* mechanism.

### Non-Partisan Legislature

In the *non-partisan legislature* mechanism, we assume that when the council rejects the mayor's office, a reversal stage starts with the selection of one councilor randomly. Suppose that there were $k$ rejections, and we are at the $k+1$-th stage in the game. A given councilor will accept the proposer's offer if, and only if, accept the offer is better than wait until the next stage. If the offer is $x_i$, then:

\[
x_i \geq \dfrac{1}{N}\left[ \delta^{k+1}R - \dfrac{N}{2}x_i \right] + \left( 1 - \dfrac{1}{N} \right)\left( \dfrac{1}{2} \right)\left[ x_i \right]
\]

In the left-hand side, we place the offer. In the right-hand side, there are two components. The first is the amount the councilor $i$ gets when she is the proposer. It is equal to the budget in the next round minus the offers that she has to make to convince half of the councilors. The second part is the gains if she happens to reject but still receive an offer in the following round. Note that the offer she makes when she is the proposer is the same as the offer she wants to receive. This because the city councilors are exchangeable, and the solution is symmetric for all councilors receiving an offer (this means that we could have dropped the $i$ in the solution). After some algebra, the offer $x_i$ has to be greater than or equal to:

\[
x_i \geq \dfrac{2 \delta^{k+1}R}{2N + 1} \equiv \underline{x}(k)
\]

The proposer always offer the minimum possible to get the proposal approved. In the case, the offer at any given stage $k$ is going to be equal to $\underline{x}(k)$.

Proceeding backwards, at the mayor's proposal stage $k=0$ and the mayor is going to offer $\underline{x}(0, N) = \dfrac{2 \delta R}{2N + 1}$ to half of the councilors. In this context, the cost of govern is equal to:

\[
C_G(N) \ = \ \dfrac{N}{2N + 1}(\delta R)
\]

\begin{proposition}
In the *non-partisan legislature* mechanism, the costs of govern increase as $N$ increases.
\end{proposition}

The differences in costs when there are $N+1$ versus when there are $N$ legislators are equal to:

\[
\begin{aligned}
C_G(N+1) - C_G(N) \ = \ & \dfrac{N+1}{2(N+1) + 1}(\delta R) - \dfrac{N}{2N + 1}(\delta R) \\
 = \ & \dfrac{\delta R}{(2N+3)(2N+1)}  \\
> \ &0
\end{aligned}
\]

Therefore, the costs always increase in $N$.

@@

\begin{corollary}
In the *non-partisan legislature*, public goods provision decreases as the legislature size increases.
\end{corollary}

This result is crucial to understand the role of the politics in our model. If the costs of govern increase with legislature size, then any increase in public goods provision associated with legislature size has come from a different mechanism. In our case, we show that there is a *political* reason for increasing the public goods provision.

### Partisan Legislature

We now look into the other extreme, which is a fully politicized legislature. In this case, when the mayor implements the policy, it generates a value of $p$. A legislator aligned with the mayor receives a benefit $p$ while an opposition legislator receives a value of $-p$. Moreover, as in the case of the non-partisan legislature, the mayor can also distribute $x_i$ private goods for the legislator $i$. In this case, when the policy is implemented, utility of an aligned legislator is equals to $x_i + p$. When the legislator is not aligned with the mayor, her utility is equals to $x_i - p$. If the policy is not implemented, then we assume a reversal policy of zero in all choice vectors $g = x_i = p = 0$.

A legislator aligned with the mayor will always support the mayor's policy, regardless of any private goods. This because as $p > 0$, the councilor always receive positive benefits from the policy. However, a politician not aligned with the mayor will require a compensation if her vote is needed. In this case, if the mayor needs the support of an opposition politician to pass her policy proposal, then she will need to compensate her for the policy costs. In this case:

\[
x_i \geq p
\]

Optimality dictates that the mayor will offer $x_i = p$ to an opposition legislator. Let $\gamma$ be the ex-ante probability of electing a politician aligned with the mayor. Then, the costs of govern when the legislature size is $N$ is equal to the expected number of politicians in the opposition that the mayor will have to compensate:

\[
C_G(N) \ = \ p\left(\dfrac{N}{2} - \gamma N\right)
\]

\begin{proposition}
In the *partisan legislature* mechanism, if the chance of electing a mayor-aligned politician is greater than one-half, then the costs of govern decrease as $N$ increases.
\end{proposition}

@@

The differences in costs when there are $N+1$ versus when there are $N$ legislators are equal to:

\[
\begin{aligned}
C_G(N+1) - C_G(N) \ = \ & p\left(\dfrac{N+1}{2} - \gamma (N+1)\right) - p\left(\dfrac{N}{2} - \gamma N\right) \\
 = \ & p\left( \dfrac{1}{2} - \gamma  \right)
\end{aligned}
\]

Therefore, when $\gamma < \dfrac{1}{2}$, the costs of govern increase. Otherwise, the costs decrease.

@@

\begin{corollary}
In the *partisan legislature*, if the chance of electing a mayor-aligned politician is greater than $1/2$, then public goods provision increase as the legislature size increases.
\end{corollary}

This result shows that when the chances of electing an aligned politician is sufficiently high, then the costs of govern decrease. As a response, the public goods provision (and also rents), increase. The rate distribution between these two vectors will be determined by the marginal change in the utility from rents and the utility from reelection. In any case, the amount allocated for both increase.

### Hybrid Partisan and Non-partisan Legislatures

The hybrid mechanism combines both partisan and non-partisan motivations. In the hybrid mechanism, a city councilor aligned with the government favors the mayor's offer if:

\[
x_i \geq \dfrac{2\delta R}{2N+1} - p
\]

An opposition politician, on the other hand, favors the mayor's offer when:

\[
x_i \geq \dfrac{2\delta R}{2N+1} + p
\]

Inspecting these equations we can decompose these costs into costs in terms of rents and costs (or benefits) from political alignment. In this context, the cost of govern will be a weighted average between the expected payments for councilors that belong to the mayor's coalition, versus the expected payments for the councilors that belongs to the opposition. The chance of a opposition member be elected is equal to $\gamma$, and taking the weighted averages, the costs become:

\[
C_G(N) \ = \ \dfrac{N}{2}\left( \dfrac{2\delta R}{2N+1} - p \right) + \left(\dfrac{N}{2} - \gamma N \right)p
\]

\begin{proposition}
In the *hybrid legislature*, if $\gamma \leq \dfrac{1}{p}\left[ \dfrac{1}{(2N+1)(2N+3)} \right]$, then the costs of govern increase as the legislature size increases. Otherwise, the costs decrease when the legislature size increases.
\end{proposition}

@@

The differences in costs when there are $N+1$ versus when there are $N$ legislators are equal to:

\[
\begin{aligned}
C_G(N+1) - C_G(N) \ = \ & \dfrac{N+1}{2}\left( \dfrac{2\delta R}{2(N+1)+1} - p \right) + \left(\dfrac{N+1}{2} - \gamma (N+1) \right)p \\
& - \dfrac{N}{2}\left( \dfrac{2\delta R}{2N+1} - p \right) - \left(\dfrac{N}{2} - \gamma N \right)p \\
 = \ & \dfrac{\delta R}{(2N+3)(2N+1)}\left[ 1-\gamma (2N+1)(2N+3)p \right]
\end{aligned}
\]

The conditions for the differences in governing costs be decrease when the  legislature size increases is:

\[
\gamma \geq \dfrac{1}{p}\left[\dfrac{1}{(2N+3)(2N+1)}\right] \equiv \underline{\gamma}
\]

@@

\begin{corollary}
In the *hybrid legislature* mechanism, public goods provision decreases as the legislature size increases.
\end{corollary}

Also, note that the threshold $\gamma$ for the probability of electing a politician aligned with the government is decreasing in both the political outcome $p$ and the size of the legislature $N$. This means that more political benefits from policy and larger legislatures makes it easier to satisfy the electoral constraint.

### Main Hypotheses

Taking stocks, the models provide two empirically testable hypotheses. Consider a legislature that previously had $N$ legislators, but now needs to increase its legislature by $N+1$. Then:

@@
